Transform and Tell:\\The News Image Caption Writer That Remembers Rare Names

### 2019-11-14 paper/section_3_model.tex
Merge changes to model from Alex

@@

Our proposed model takes three inputs: the image, the faces, and article text.
Each of these inputs first goes through an encoder to give us a vector
representation.

\subsubsection{Image Embedding}

For the image, we feed it through a pretrained ResNet-152 model
\cite{He2016ResNet} and use the output of the final block, just before the
pooling layer, as the image embedding $\bX_I \in \mathbb{R}^{2048 \times 49}$.
The embedding forms a 7 by 7 block, allowing the transformer to attend to 49
different patches in the image.

\subsubsection{Article Embedding}

For the article, we use RoBERTa \cite{Liu2019RoBERTaAR} to encode the text.
RoBERTa, a more carefully trained BERT \cite{Devlin2019BERT}, is a language
representation model that provides pretrained contextual embeddings for text.
Unlike GloVe \cite{Pennington2014Glove} and word2vec
\cite{Mikolov2013DistributedRO} embeddings, where each word has exactly one
representation, the bidirectionality and the attention mechanism in the
transformer allow a word to have different vector representations depending on
the surrounding context.

@@

Let $\bh_t \in \mathbb{R}^{1024}$ be the current hidden state when decoding the
$t$th token. We divide the embedding dimension into 16 attention heads, each
with size 64. For each head $i \in \{1, 2, ..., 16\}$, we first do a linear
projection of $\bh_t$ and the image embedding $\bX_I$ into a query $\bq_{It}
\in \mathbb{R}^{64}$, key $\bK_I \in \mathbb{R}^{64 \times 49}$, and value
$\bV_I \in \mathbb{R}^{64 \times 49}$:
\begin{IEEEeqnarray*}{lCl}
   \bq_{Iit} &=& \bW_{Iiq} \, \bh_t \\
   \bK_{Ii} &=& \bW_{Iik} \, \bX_I \\
   \bV_{Ii} &=& \bW_{Iiv} \, \bX_I
\end{IEEEeqnarray*}
Then the attended image for each head $\bx'_{Ii} \in \mathbb{R}^{64}$ is the
weighted sum of the values, where the weights are obtained from the dot product
between the query and key:
\begin{IEEEeqnarray*}{lCl}
   \bx'_{Ii} &=& \text{softmax}\left(\bK_{Ii}^T \bq_{Iit} \right) \bV_{Ii}
\end{IEEEeqnarray*}
The attention from each head is then concatenated into $\bx'_{I} \in
\mathbb{R}^{1024}$:
\begin{IEEEeqnarray*}{lCl}
   \bx'_{I} &=& [\tx_{I1}, \tx_{I2}, ..., \tx_{I16}]
\end{IEEEeqnarray*}
and the overall image attention $\tx_{I} \in \mathbb{R}^{1024}$ is obtained
after adding a residual connection and layer normalization:
\begin{IEEEeqnarray*}{lCl}
   \tx_{I} &=& \text{LayerNorm}(\bh_t + \bx'_{I})
\end{IEEEeqnarray*}
We use the same attention mechanism (with different weight matrices) to obtain
the attended article $\tx_T$ and the attended face $\tx_F$. These three are
then concatenated and fed through a feedforward layer:
\begin{IEEEeqnarray*}{lCl}
   \tx_C &=& [\tx_I, \tx_T, \tx_F] \\
   \tx_R &=& \bW_C \, \tx_C + \bb_C \\
   \tx_D &=& \text{ReLU}(\bW_R \, \tx_R + \bb_R )\\
   \tx_E &=& \text{LayerNorm}(\tx_R + \bW_D \, \tx_D + \bb_D)
\end{IEEEeqnarray*}
The final output $\tx_E \in \mathbb{R}^{1024}$ is used as the input to the
next transformer block.

@@

Instead of using the standard self-attention module as in current
state-of-the-art GPT-2 architecture \cite{Radford2019LanguageMA}, we find that
dynamic convolutions \cite{Wu2018PayLA} are more efficient to train, especially
when there is limited GPU compute power.

Given an input $\bz_t \in \mathbb{R}^{1024}$, which can either be the input
embedding of the previous token, or the output from the previous transformer
block, we first use a feedforward layer with a gated linear unit (GLU)
\cite{Dauphin2017GLU} as the activation function:
\begin{IEEEeqnarray*}{lCl}
   \bz_t' &=& \text{GLU}(\bW_Z \, \bz_t + \bb_Z)
\end{IEEEeqnarray*}

>>>>

Conceptually our model can be broken into two parts: encoding and decoding. The
encoding part consists of a set of domain specific encoders for producing high
level vector representations of images, faces and article text. The output of
each encoder is a, potentially arbitrary, length set of fixed size vectors that
represent the input. The decoder sequentially generates captions at the
sub-word level by applying
multi-headed attention over the sets of vectors from the encoders, and over a
representation of the previously generated sub-word units. In
practice there are a number details which allow us to train this large
multi-faceted model on a single machine and achieve state-of-the-art
performance. We have included some these details where appropriate and
collected together those that do not fit elsewhere into
Section~\ref{ssec:bag_of_tricks}.

@@

Our proposed model takes three types of inputs: image, faces, and article text.
Each of these inputs is encoded into a set of vectors by domain specific
encoders pre-trained on data from the matching domain.

\subsubsection{Image Encoder}

A high level image representation is obtained with a
ResNet-152~\cite{He2016ResNet} model pre-trained on ImageNet. We use the output
of the final block before the
pooling layer as the image representation.
This is set of 49 different vectors $x_{I,i} \in \mathbb{R}^{2048}$ where each
vector corresponds to a separate image patch after the image is divided into
equal size 7 by 7 patches. Using this representation $\bx_I = \{x_{I,i} \in
\mathbb{R}^{2048}\}_{i=0}^{49}$ allows the decoder to attend different regions
in the
image -- a modelling choice that has proven useful in other image captioning
tasks~\cite{Xu2015ShowAA}.

\subsubsection{Face Encoder}

We use MTCNN \cite{Zhang2016JointFD} to detect face bounding boxes in the
image. We then select the largest $M$ faces. A vector representation of each
face is obtained by passing the bounding boxes to
FaceNet~\cite{Schroff2015FaceNetAU}, which was pre-trained on the VGGFace2
dataset~\cite{Cao2017VGGFace2AD}. The resulting set of face vectors for each
image is $\bx_{F} = \{\bx_{F,i} \in \mathbb{R}^{512}\}_{i=0}^M$.

Even though the faces are extracted from the image it is useful to consider
them as an input domain that is separate to the image. This is because
specialised models are needed to make full use of them.

%them to be a separate
%input domain since face bounding boxes are extracted as a pre-processing step
%and we use a specialised model to encode features from each bounding box.

\eat{We then select the top $M$ faces and feed them through a
FaceNet model \cite{Schroff2015FaceNetAU}, pretrained on the VGGFace2 dataset
\cite{Cao2017VGGFace2AD}, to obtain a face embedding $\bx_{F} =
\{\bx_{F,i} \in \mathbb{R}^{512}\}_{i=0}^M$.}

\subsubsection{Article Encoder}

To encode the article text we use RoBERTa~\cite{Liu2019RoBERTaAR} which is a
recent improvement over the popular BERT~\cite{Devlin2019BERT} model.
RoBERTa is a language representation model that provides pretrained contextual
embeddings for text. It consists of 24 layers of bidirectional transformer
blocks.
% It is a more carefully trained BERT \cite{Devlin2019BERT} model.

Unlike GloVe \cite{Pennington2014Glove} and word2vec
\cite{Mikolov2013DistributedRO} embeddings, where each word has exactly one
representation, the bidirectionality and the attention mechanism in the
transformer allow a word to have different vector representations depending on
the surrounding context.

@@

Let $\bh_t \in \mathbb{R}^{1024}$ be the current hidden state when decoding the
$t$th token. We use 16 attention heads, each with size 64. For each head $i \in
\{1, 2, ..., 16\}$, we first do a linear projection of $\bh_t$ and the image
embedding $\bX_I$ into a query $\bq_{It} \in \mathbb{R}^{64}$, key $\bK_I \in
\mathbb{R}^{64 \times 49}$, and value $\bV_I \in \mathbb{R}^{64 \times 49}$:
\begin{IEEEeqnarray*}{lCl}
   \bq_{Iit} &=& \bW_{Iiq} \, \bh_t \\
   \bK_{Ii} &=& \bW_{Iik} \, \bX_I \\
   \bV_{Ii} &=& \bW_{Iiv} \, \bX_I
\end{IEEEeqnarray*}
Then the attended image for each head $\bx'_{Ii} \in \mathbb{R}^{64}$ is the
weighted sum of the values, where the weights are obtained from the dot product
between the query and key:
\begin{IEEEeqnarray*}{lCl}
   \bx'_{Ii} &=& \text{softmax}\left(\bK_{Ii}^T \bq_{Iit} \right) \bV_{Ii}
\end{IEEEeqnarray*}
The attention from each head is then concatenated into $\bx'_{I} \in
\mathbb{R}^{1024}$:
\begin{IEEEeqnarray*}{lCl}
   \bx'_{I} &=& [\tx_{I1}, \tx_{I2}, ..., \tx_{I16}]
\end{IEEEeqnarray*}
and the overall image attention $\tx_{I} \in \mathbb{R}^{1024}$ is obtained
after adding a residual connection and layer normalization:
\begin{IEEEeqnarray*}{lCl}
   \tx_{I} &=& \text{LayerNorm}(\bh_t + \bx'_{I})
\end{IEEEeqnarray*}
We use the same attention mechanism (with different weight matrices) to obtain
the attended article $\tx_T$ and the attended face $\tx_F$. These three are
then concatenated and fed through a feedforward layer:
\begin{IEEEeqnarray*}{lCl}
   \tx_C &=& [\tx_I, \tx_T, \tx_F] \\
   \tx_R &=& \bW_C \, \tx_C + \bb_C \\
   \tx_D &=& \text{ReLU}(\bW_R \, \tx_R + \bb_R )\\
   \tx_E &=& \text{LayerNorm}(\tx_R + \bW_D \, \tx_D + \bb_D)
\end{IEEEeqnarray*}
The final output $\tx_E \in \mathbb{R}^{1024}$ is used as the input to the
next transformer block.

@@

Instead of using the standard self-attention module as in current
state-of-the-art GPT-2 architecture \cite{Radford2019LanguageMA}, we find that
dynamic convolutions \cite{Wu2018PayLA} are more efficient to train. Suppose
when decoding the $t$th token, we have an input $\bz_t \in \mathbb{R}^{1024}$,
which can either be the input embedding of the previous token, or the output
from the previous transformer block. Given kernel size $K$, we first do a
projection of the last $K$ steps using a feedforward layer with a gated linear
unit (GLU) \cite{Dauphin2017GLU} as the activation function:
\begin{IEEEeqnarray*}{lCl}
   \bz_{t-j}' &=& \text{GLU}(\bW_Z \, \bz_{t-j} + \bb_Z) \qquad \forall j \in \{0,1,...,K-1\}
\end{IEEEeqnarray*}
the output of the dynamic convolution is simply the weighted sum of these
projected values:
\begin{IEEEeqnarray*}{lCl}
   \tz_t &=& \sum_0^{K-1}
\end{IEEEeqnarray*}

We use 16 attention heads, each with size 64. For each head $i \in
\{1, 2, ..., 16\}$, we first do a linear projection to get the kernel weight
\begin{IEEEeqnarray*}{lCl}
   \bW \, \bz_t'
\end{IEEEeqnarray*}
